---
title: 'L4 AA Analysis'
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

##AA analysis
goals of analysis
look at metals at different sites (compare means?)
instrumentation - compare how ICPMS and AA differ for Cr

loading data
```{r}
AA <- read.csv("~/CHEM313_L4_SoilLab/data/tidy_AA.csv")
```
defining lists (probably won't use)
```{r}
sample_sites <- unique(filter(AA, site!="MB", site!="")$site)
```
making cal curve
```{r}
cal <- AA %>%
  filter(type == "CalStd" | type == "CalStd2" | type == "CalStd4") %>%
  select(concentration, mean_abs, percent_rsd)
  
model <- lm(cal$mean_abs ~ cal$concentration)
slope <- model$coefficients[2]
intercept <- model$coefficients[1]
slope_std <- summary(model)$coefficients[2,2]
intercept_std <- summary(model)$coefficients[1,2]
  
plot(cal$mean_abs ~ cal$concentration,
     xlab = "Concentration of Cr (ppm)",
     ylab = "Absorbance") +
  abline(model, col="blue") +
  title("Calibration for Cr")

AA_cal <- tibble(slope, slope_std, intercept, intercept_std)
AA_cal
#clean up
rm(cal, slope, slope_std, intercept, intercept_std, model)
```

fitting samples to cal curve
```{r}
AA_sample <- filter(AA, site != "")
m <- AA_cal$slope
b <- AA_cal$intercept
y <- AA_sample$mean_abs
b_e <- AA_cal$intercept_std
m_e <- AA_cal$slope_std
x <- (y-b)/m
RSD <- AA_sample$percent_rsd/100
ABS <- AA_sample$mean_abs
#error propagation
yb <- ABS-b
e_yb <- sqrt((y*RSD)^2 + (b_e)^2) #error in y-b in calibration
e_x <- x*sqrt((e_yb/yb)^2+(m_e/m)^2) #error in x from calibration

AA_concentration_data <- tibble(sample_key = AA_sample$sample_key,
                                         analyst = AA_sample$analyst,
                                         site = AA_sample$site,
                                         conc_dil = x,
                                         conc_dil_error = e_x)

AA_concentration_data
```

method blank
```{r}
MB <- filter(AA_concentration_data, site == "MB")
MB_conc <- mean(MB$conc_dil)
MB_conc_error <- sd(MB$conc_dil)

MB
MB_conc_error
```

adjusting data for method blank
```{r}
AA_forblank <- filter(AA_concentration_data, site != "MB")
conc_dil_blanked <- AA_forblank$conc_dil - MB_conc
conc_dil_blanked_error <- sqrt((AA_forblank$conc_dil_error)^2 + (MB_conc_error)^2)
AA_blank <- AA_forblank %>% mutate(conc_dil_blanked, conc_dil_blanked_error)

AA_blank
```

dilution correction and error propagation

error propagation and correcting dilutions
```{r}
#defining dilution factors and measurement errors
vol_e <- 1
mass_e <- .001
dil_1010_e <- sqrt(1^2 + 10^2)
dil_e <- sqrt((dil_1010_e/1010)^2+(1/10)^2)

#correct for dilutions
AA_corrected <- merge(AA, AA_blank) %>%
  unique() %>%
  mutate(conc_blanked = conc_dil_blanked*(total_volume/1000)/(mass_of_soil/1000),
    conc_blanked_error = conc_blanked*sqrt((conc_dil_blanked_error/conc_dil_blanked)^2+
                                            (dil_e/101)^2 +
                                            (mass_e/mass_of_soil)^2 +
                                            (vol_e/total_volume)^2),
    conc_unblanked = conc_dil*(total_volume/1000)/(mass_of_soil/1000),
    conc_unblanked_error = conc_unblanked*sqrt((conc_dil_error/conc_dil)^2 +
                                                (dil_e/101)^2 +
                                                (mass_e/mass_of_soil)^2 +
                                                (vol_e/total_volume)^2)) %>%
  select(-concentration, #remove unnecessary columns
         -type,
         -X,
         -mass_of_soil,
         -total_volume,
         -mean_abs,
         -percent_rsd,
         -conc_dil_blanked,
         -conc_dil_blanked_error,
         -conc_dil,
         -conc_dil_error)

AA_corrected

rm(list = ls()[!ls() %in% c("AA", "AA_blank", "AA_cal", "AA_corrected", "MB")])
```

DETECTION/QUANTIFICATION LIMIT OF FAAS
```{r}
lod <- 3*sd(MB$conc_dil)/AA_cal$slope
lod

qod <- 10*sd(MB$conc_dil)/AA_cal$slope
qod
```